Draft - Enhancing Forecast Reconciliation: A Study of Covariance Estimation

### 2025-04-24 literature/paper_draft.Rmd
writing simulation and literature review

---
title: "Outline - Covariance Estimation for Hierarchical/Grouped Time Series Using the NOVELIST Estimator"
author: "Vincent Su"
date: "`r Sys.Date()`"
output: 
  pdf_document
header-includes:
  - \usepackage{bm}
  - \usepackage{titlesec}
  - \titlespacing{\section}{0pt}{30pt plus 12pt minus 1pt}{0pt plus 1pt minus 1pt}
# bibliography: bibliography.bib
---

@@

A collection of time series connected via a set of linear constraints is known as hierarchical time series. Forecasting these structures is challenging because the individual forecasts do not satisfy the linear constraints present in the data. To mitigate this issue, various forecast reconciliation approaches have been proposed in the literature, where the individual forecasts are adjusted by minimizing the mean squared reconciliation errors. Among these, the MinT (Minimum Trace) approach is widely used. However, this method requires an estimate of the covariance matrix of the base forecast errors. Although the sample covariance matrix is a natural choice, its performance is greatly impacted by the high-dimensional nature of hierarchical time series encountered in practice. Most published research uses shrinkage-type estimators that shrink the sample covariance matrix toward a diagonal matrix.

In this project, we aim to assess the forecasting performance of MinT when different covariance estimators are used, with a primary focus on the NOVELIST (NOVEL Integration of the Sample and Thresholded Covariance) estimator, which shrinks the sample covariance toward its thresholded version. A thorough literature review will also identify other high-dimensional covariance estimators relevant to this project.

@@

-   **Problem Statement:**\
    The sample covariance matrix, although natural, suffers in high-dimensional settings. Especially when the number of series $p$ is huge and larger than the time dimension $T$, the sample covariance matrix is non-positive definite (rank T if p\>T).

@@

-   **Research Aim:**\
    This paper assesses the reconciled forecasting performance of MinT approach using various covariance estimators, with a focus on the NOVELIST estimator.

-   **Paper Outline:**\
    The paper is structured as follows:

@@

-   Brief history of reconciliation methods (bottom-up, top-down, GLS, MinT).
-   Statistical challenges and practical applications of MinT.

@@

-   $W_h$ is challenging to estimate, when $h$ \> 1, we need alternative estimates.

>>>>

---
title: "Draft - Enhancing Forecast Reconciliation: A Study of Covariance Estimation"
author: "Vincent Su"
date: "`r Sys.Date()`"
output: 
  pdf_document
header-includes:
  - \usepackage{bm}
  - \usepackage{titlesec}
  - \titlespacing{\section}{0pt}{30pt plus 12pt minus 1pt}{0pt plus 1pt minus 1pt}
# bibliography: bibliography.bib
---

@@

A collection of time series connected via a set of linear constraints is known as hierarchical time series. Forecasting these structures is challenging because the individual forecasts do not satisfy the linear constraints present in the data. To mitigate this issue, various forecast reconciliation approaches have been proposed in the literature, where the individual forecasts are adjusted by minimizing the mean squared reconciliation errors. Among these, the MinT (Minimum Trace) approach is widely used, and this method requires an estimate of the covariance matrix of the base forecast errors.

<!-- Although the sample covariance matrix is a natural choice, its performance is greatly impacted by the high-dimensional nature of hierarchical time series encountered in practice. Most published research uses shrinkage-type estimators that shrink the sample covariance matrix toward a diagonal matrix. -->

In this project, we aim to assess the forecasting performance of MinT when different covariance estimators are used, and compare with the proposed estimators from published paper. with a primary focus on the NOVELIST (NOVEL Integration of the Sample and Thresholded Covariance) estimator, which shrinks the sample covariance toward its thresholded version. A thorough literature review will also identify other high-dimensional covariance estimators relevant to this project.

@@

-   **Problem Statement:**

    The sample covariance matrix, although natural, suffers in high-dimensional settings. Especially when the number of series $p$ is huge and larger than the time dimension $T$, the sample covariance matrix is non-positive definite (rank T if p\>T).

@@

-   **Research Aim:**

    This paper assesses the reconciled forecasting performance of MinT approach using various covariance estimators, with a focus on the NOVELIST estimator.

-   **Paper Outline:**

    The paper is structured as follows:

@@

Forecast reconciliation converts a collection of independent base forecasts into a set of coherent forecasts that respect the linear constraints defining a hierarchical or grouped time-series system. Early work focused on heuristic single-level strategies, including bottom-up, top-down, and middle-out (...), each of which exploits only part of the information in the hierarchy and can induce bias or high variance.

- Cite the single level approach
- Athanasopoulos et al., 2024

Hyndman et al. (2011) first showed that all single-level methods can be written as $\tilde{y} = SG\hat{y}$, where $S$ is the summing matrix and $G$ is a matrix that maps base forecasts $\hat{y}$ to into the bottom level. Treating reconciliation as a GLS regression problem, Hyndman et al. (2011) found that it yields a solution for $G$, but the required covariance matrix of reconcilation error is not identifiable in practice (Wickramasuriya et al., 2019).

- (Talk more about how others transform it to OLS, WLS,..)
- Di Fonzo and Marini (2011)
- Athanasopoulos et al. (2009)

Wickramasuriya, Athanasopoulos & Hyndman (2019) reframed the problem by taking an optimisation approach rather than the regression. They formulated the problem as minimising the variances of all reconciled forecasts, which happens to be equivalent to minimising the trace of the covariance matrix (sum of the diagonal elements). This is known as the Minimum Trace (MinT) reconciliation method. The MinT solution is given by $G_h = (S'W_h^{-1}S)^{-1}S'W_h^{-1}$, and $W_h$ is the covariance matrix of the h-step-ahead base forecast errors.

The MinT approach is an algebraical generalisation of the GLS, and the OLS and WLS methods are special cases of MinT when $W_h$ is a diagonal or identity matrix, respectively. However, the MinT solution hinges on a reliable estimate of the h-step-ahead base forecast error covariance $W_h$. In high-dimensional setting, the usual sample covariance matrix is unstable, thus we need alternative covariance estimators.

- Structural Scaling, based only on the struc- ture of the hierarchy (Athanasopoulos et al., 2017)
- Shrinkage estimators (Schäfer & Strimmer, 2005; Ledoit & Wolf, 2004)

Empirical evaluations have demonstrated that MinT with an appropriate covariance estimate often outperforms earlier methods in both simulation and real data studies.

<!-- Wickramasuriya et al. (2019) adopt the diagonal-target Ledoit–Wolf approach and show analytically that any scalar multiple $k_h$ from $W_h = k_hW_1$ suffices for point-forecast reconciliation.  However, simulation and empirical evidence suggest that ignoring cross-series error correlation can leave accuracy untapped, especially in highly-coupled systems (energy, retail SKUs).  Their subsequent work imposes non-negativity constraints via quadratic programming citeturn0file2 and studies bias-robust variants (MinT-U). -->

<!-- Empirical findings to date   -->
<!-- Athanasopoulos et al. (2024) synthesise the accumulated evidence: MinT dominates single-level baselines in most studies; WLS or diagonal-shrinkage MinT is the current industry default; but gains plateau when off-diagonals are weak citeturn0file9.  Only a handful of papers explore richer covariance structures (e.g., block or factor shrinkage), and none benchmark them systematically against NOVELIST-type estimators. -->

@@

-   $W_h$ is challenging to estimate, when $h>1$, we need alternative estimates.
